---
title: "Web APIs"
author: "Pavan Akula"
date: "March 31, 2017"
output:
  html_document: default
  pdf_document: default
---

###DATA607-Week09 - Assignment

The purpose of this assignment is to connect to NYTimes.com website, extract information and store into R data frame for analysis. According to NYTimes.com developer [FAQ](http://developer.nytimes.com/faq) more API's may be released over time. Following are the steps:

1. Get an API Key from the website http://developer.nytimes.com/signup.
2. Select an API to extract data.
3. Data extraction and storage.
4. Data analysis.

Libraries used
```{r, message=F}
library(urltools)
library(jsonlite)
library(knitr)
library(dplyr)
library(stringr)
library(ggplot2)
```

####Get an API Key from the website
Open [NYTimes Create a New API Key](http://developer.nytimes.com/signup) web page and key in all the required information and click on [Create API Key.](http://developer.nytimes.com/signup) NYTimes will send an alpha-numeric key to the email address.

####Select an API to extract data
Once API Key is acquired, validate it by visiting web page http://developer.nytimes.com/. Pick any topic of your choice and follow the instructions.

For the purpose of the assignment, I have picked two API's "Most Popular API" and "Article Search API".

#####Most Popular API
With the Most Popular API, you can get links and metadata for the blog posts and articles that are most frequently e-mailed, shared and viewed by NYTimes.com readers. Parameter accepted by this API are "section" and "time-period" Additional information can be found at http://developer.nytimes.com/most_popular_api_v2.json#/README.

Allowed values for the section:

- Arts
- Automobiles
- Blogs
- Books
- Business Day
- Education
- Fashion & Style
- Food
- Health
- Job Market
- Magazine
- membercenter
- Movies
- Multimedia
- N.Y.%20%2F%20Region
- NYT Now
- Obituaries
- Open
- Opinion
- Public Editor
- Real Estate
- Science
- Sports
- Style
- Sunday Review
- T Magazine
- Technology
- The Upshot
- Theater
- Times Insider
- Today's Paper
- Travel
- U.S.
- World
- Your Money
- all-sections

Allowed values for time-period (represents days):

- 1
- 7
- 30

```{r, echo=T}
#URL formation
#Base URL
nyt.base.url <- "https://api.nytimes.com/svc/mostpopular/v2/"

#Second part, path changes based on path, section, and time-period
#{mostemailed | mostshared | mostviewed}

#Possible paths
nyt.path <- data.frame(DisplayText=NA,UrlPath=NA)
nyt.path <- rbind(nyt.path, c("Most Emailed","mostemailed"))
nyt.path <- rbind(nyt.path, c("Most Shared","mostshared"))
nyt.path <- rbind(nyt.path, c("Most Viewed","mostviewed"))
nyt.path <- na.omit(nyt.path)

#Possible sections
nyt.sec<- c("Arts","Automobiles","Blogs","Books","Business Day","Education","Fashion & Style","Food","Health","Job Market","Magazine","membercenter","Movies","Multimedia","N.Y.%20%2F%20Region","NYT Now","Obituaries","Open","Opinion","Public Editor","Real Estate","Science","Sports","Style","Sunday Review","T Magazine","Technology","The Upshot","Theater","Times Insider","Today's Paper","Travel","U.S.","World","Your Money","all-sections")

#Possible time-period
nyt.time <- data.frame(DisplayText=NA,TimePath=NA)
nyt.time <- rbind(nyt.time, c("1 day","1.json"))
nyt.time <- rbind(nyt.time, c("7 days","7.json"))
nyt.time <- rbind(nyt.time, c("30 days","30.json"))
nyt.time <- na.omit(nyt.time)

nyt.path.data.frame <- data.frame(Path=NA,NYTUrl=NA)
#loop through path
for (p in 1:nrow(nyt.path)){
  #loop through sections
  for(s in 1:length(nyt.sec)){
    #loop through time
    for (t in 1:nrow(nyt.time)){
      
      # Generate complete URL
      nyt.complete.url <- paste0(nyt.base.url, nyt.path$UrlPath[p], "/", nyt.sec[s], "/", nyt.time$TimePath[t])
      
      #Attach API Key
      nyt.complete.url <- param_set(nyt.complete.url, key = "api-key", value = "a8a7bea8b95f4ef986e504ec99287013")
      
      nyt.complete.url <- URLencode(nyt.complete.url) #Converts spaces and special characters to URL format
      
      #Save the URL
      nyt.url.info <- paste0(nyt.path$DisplayText[p]," by ", nyt.sec[s], " in ", nyt.time$DisplayText[t])
      nyt.path.data.frame <- rbind(nyt.path.data.frame, c(nyt.url.info, nyt.complete.url))
    }
  }
}

nyt.path.data.frame <- na.omit(nyt.path.data.frame)
row.names(nyt.path.data.frame) <- NULL

set.seed(125)
nyt.path.data.sample <- nyt.path.data.frame %>% sample_n(., 10) 
row.names(nyt.path.data.sample) <- NULL

kable(nyt.path.data.sample, format="pandoc", align="l", col.names = c("Path","NYTUrl"), row.names = NA,caption = "Sample list of URL's for *_NYTimes.com Most Popular API_*")

```

#####Article Search API
You can get useful results from the Article Search API with a simple query. But you can also use filters and facets to extend the functionality of the API. Additional information on filters and sorting can be found at http://developer.nytimes.com/article_search_v2.json#/Documentation/GET/articlesearch.json.

```{r, echo=T}
nyt.path.data.frame <- data.frame(Path=NA,NYTUrl=NA)
#URL formation
#Base URL
nyt.base.url <- "https://api.nytimes.com/svc/search/v2/articlesearch.json"

#Attach API Key
nyt.complete.url <- param_set(nyt.base.url, key = "api-key", value = "a8a7bea8b95f4ef986e504ec99287013")
nyt.complete.url <- URLencode(nyt.complete.url)
nyt.url.info <- "Generic Article Search"
nyt.path.data.frame <- rbind(nyt.path.data.frame, c(nyt.url.info, nyt.complete.url))

#Search query term. Search is performed on the article body, headline and byline looking for word apple.
nyt.complete.url <- param_set(nyt.base.url, key = "q", value = "Apple")
nyt.complete.url <- param_set(nyt.complete.url, key = "api-key", value = "a8a7bea8b95f4ef986e504ec99287013")
nyt.complete.url <- URLencode(nyt.complete.url)
nyt.url.info <- "Query Article Search with `q`"
nyt.path.data.frame <- rbind(nyt.path.data.frame, c(nyt.url.info, nyt.complete.url))

#The filter query results, search for word "Apple" and filter by "Court"
nyt.complete.url <- param_set(nyt.base.url, key = "q", value = "Apple")
nyt.complete.url <- param_set(nyt.complete.url, key = "fq", value = "Court")
nyt.complete.url <- param_set(nyt.complete.url, key = "api-key", value = "a8a7bea8b95f4ef986e504ec99287013")
nyt.complete.url <- URLencode(nyt.complete.url)
nyt.url.info <- "Filter Query results with `fq`"
nyt.path.data.frame <- rbind(nyt.path.data.frame, c(nyt.url.info, nyt.complete.url))

#Add date conditions Format: YYYYMMDD
nyt.complete.url <- param_set(nyt.base.url, key = "q", value = "Apple")
nyt.complete.url <- param_set(nyt.complete.url, key = "fq", value = "Court")
nyt.complete.url <- param_set(nyt.complete.url, key = "begin_date", value = "20160101")
nyt.complete.url <- param_set(nyt.complete.url, key = "end_date", value = "20160131")
nyt.complete.url <- param_set(nyt.complete.url, key = "api-key", value = "a8a7bea8b95f4ef986e504ec99287013")
nyt.complete.url <- URLencode(nyt.complete.url)
nyt.url.info <- "Filter Query results with `begin_date` and `end_date` `Format: YYYYMMDD`"
nyt.path.data.frame <- rbind(nyt.path.data.frame, c(nyt.url.info, nyt.complete.url))

#Sorting data
nyt.complete.url <- param_set(nyt.base.url, key = "q", value = "Apple")
nyt.complete.url <- param_set(nyt.complete.url, key = "fq", value = "Court")
nyt.complete.url <- param_set(nyt.complete.url, key = "begin_date", value = "20160101")
nyt.complete.url <- param_set(nyt.complete.url, key = "end_date", value = "20160131")
nyt.complete.url <- param_set(nyt.complete.url, key = "sort", value = "newest")
nyt.complete.url <- param_set(nyt.complete.url, key = "api-key", value = "a8a7bea8b95f4ef986e504ec99287013")
nyt.complete.url <- URLencode(nyt.complete.url)
nyt.url.info <- "Sort Query results with `sort` `newest or oldest`"
nyt.path.data.frame <- rbind(nyt.path.data.frame, c(nyt.url.info, nyt.complete.url))

nyt.path.data.frame <- na.omit(nyt.path.data.frame)
row.names(nyt.path.data.frame) <- NULL

kable(nyt.path.data.frame, format="pandoc", align="l", col.names = c("Path","NYTUrl"), row.names = NA,caption = "Sample list of URL's for *_NYTimes.com Article Search API_*")
```

####Data extraction and storage
Following code extracts JSON data using **_Most Popular API_** and stores into R data frames namely `mostemailed.data`, `mostshared.data` and `mostviewed.data`.

```{r, echo=T}
#Extracting data using Most Popular API for World 7 days path
most<-c("mostemailed","mostshared","mostviewed")
nyt.base.url <- "https://api.nytimes.com/svc/mostpopular/v2/mostpath/World/7.json"

for (i in 1:length(most)){
  #First set of results
  nyt.complete.url <- str_replace(nyt.base.url, pattern = "mostpath", replacement = most[i])
  nyt.complete.url <- param_set(nyt.complete.url, key = "api-key", value = "a8a7bea8b95f4ef986e504ec99287013")
  nyt.complete.url <- URLencode(nyt.complete.url)
  nyt.raw.data <- fromJSON(nyt.complete.url)

  if (i==1){
    mostemailed.data <- nyt.raw.data$results
  }
  else if(i==2){
    mostshared.data <- nyt.raw.data$results
  }
  else{
    mostviewed.data <- nyt.raw.data$results
  }
}

```

Following code extracts JSON data using **_Article Search API_** and stores into R data frames namely `nyt.article.data`. Query extracts articles containing words `Apple` and `Court` between `01/01/2016` and `01/31/2016` ordered by newest.

```{r, echo=T}

nyt.base.url <- "https://api.nytimes.com/svc/search/v2/articlesearch.json"

nyt.complete.url <- param_set(nyt.base.url, key = "q", value = '"Apple"')
nyt.complete.url <- param_set(nyt.complete.url, key = "fq", value = '"Court"')
nyt.complete.url <- param_set(nyt.complete.url, key = "begin_date", value = "20160101")
nyt.complete.url <- param_set(nyt.complete.url, key = "end_date", value = "20160131")
nyt.complete.url <- param_set(nyt.complete.url, key = "sort", value = "newest")
nyt.complete.url <- param_set(nyt.complete.url, key = "api-key", value = "a8a7bea8b95f4ef986e504ec99287013")
nyt.complete.url <- URLencode(nyt.complete.url)
nyt.raw.data <- fromJSON(nyt.complete.url)

nyt.article.data <- nyt.raw.data$response$docs
row.names(nyt.article.data) <- NULL

nyt.article.data %>% select(source, snippet) %>% 
kable(format="pandoc", align="l", col.names = c("Source","Snippet"), row.names = NA,caption = "*_Sample List of Articles containing words `Apple` and `Court`_*")

```

####Data analysis
```{r, echo=T}
nyt.most.common.data <- mostemailed.data %>% 
  inner_join(mostshared.data, by = "title") %>% 
  inner_join(mostviewed.data, by = "title") %>%
  select (title, total_shares, views)

kable(nyt.most.common.data$title, format="pandoc", align="l", col.names = c("Title"), row.names = NA,caption = "List of Titles that were Most Viewed, Shared and Emailed")


popular.topic <- unlist(mostemailed.data$des_facet)
email.topic <- data.frame(topic=popular.topic, stringsAsFactors = F)

popular.topic <- unlist(mostshared.data$des_facet)
shared.topic <- data.frame(topic=popular.topic, stringsAsFactors = F)

popular.topic <- unlist(mostviewed.data$des_facet)
viewed.topic <- data.frame(topic=popular.topic, stringsAsFactors = F)

mostpopular.topic <- rbind(email.topic, shared.topic, viewed.topic)

popular.topic.summary <- mostpopular.topic %>% group_by(topic) %>% tally() %>% filter(topic != "") %>% arrange(desc(n)) %>% mutate(id=row_number()) %>%  filter(n > 1) %>% select(topic,n)

ggplot(data=popular.topic.summary, aes(x=topic, y=n)) +
  geom_bar(stat="identity") + labs(x="Topic", y="Number of Times", title = "Topics refered more than once") + coord_flip()
```


References: http://developer.nytimes.com
